---
title: "Reproducible Research - Project 1"
author: "Kevin Ho"
date: "21 May 2016"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Code for reading in the dataset and/or processing the data
```{r, cache=TRUE}
setwd("C:/Users/kxho/Documents/R-Dev/Reproducible Research")
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip","repdata-data-activity.zip")
unzip("repdata-data-activity.zip")
actData <- read.csv("activity.csv")  
```

## Histogram of the total number of steps taken each day
```{r}
stepsByDate <- tapply(actData$steps,actData$date, sum, na.rm=TRUE)
barplot(stepsByDate,names.arg = names(stepsByDate), xlab = "Date",ylab = "Steps")
```

## Mean and median number of steps taken each day

### Mean
```{r}
tapply(actData$steps,actData$date, mean, na.rm = TRUE)
```

### Median
```{r}
tapply(actData$steps,actData$date, median, na.rm = TRUE)
```

## Time series plot of the average number of steps taken
```{r}
intervalData <- tapply(actData$steps,actData$interval,mean,na.rm=TRUE)
plot(names(intervalData),intervalData, type= "l", xlab = "Time Interval", ylab = "Steps")
```


## The 5-minute interval that, on average, contains the maximum number of steps
```{r}
names(intervalData[max(intervalData)==intervalData])
```


## Code to describe and show a strategy for imputing missing data
Where the data in "NA" we can plug with the mean for that interval
```{r}
actDataPlugged <- actData
actDataPlugged$steps[is.na(actDataPlugged$steps)] <- intervalData[as.character(actDataPlugged$interval[is.na(actDataPlugged$steps)])]
```

## Histogram of the total number of steps taken each day after missing values are imputed
```{r}
stepsByDatePlugged <- tapply(actDataPlugged$steps,actDataPlugged$date, sum, na.rm=TRUE)
barplot(stepsByDatePlugged,names.arg = names(stepsByDatePlugged), xlab = "Date",ylab = "Steps")
```

### Mean
```{r}
tapply(actDataPlugged$steps,actDataPlugged$date, mean, na.rm = TRUE)
```

### Median
```{r}
tapply(actDataPlugged$steps,actDataPlugged$date, median, na.rm = TRUE)
```

By plugging the data the means and median can be calculated for more intervals. Although some values might be inflated.

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
actDataPluggedWeekday <- actDataPlugged[!(weekdays(as.POSIXlt(actDataPlugged$date)) %in% c("Sunday", "Saturday")),]
intervalDataWeekday <- tapply(actDataPluggedWeekday$steps,actDataPluggedWeekday$interval,mean,na.rm=TRUE)

actDataPluggedWeekend <- actDataPlugged[weekdays(as.POSIXlt(actDataPlugged$date)) %in% c("Sunday", "Saturday"),]
intervalDataWeekend <- tapply(actDataPluggedWeekend$steps,actDataPluggedWeekend$interval,mean,na.rm=TRUE)

par(mfrow=c(1,2))
plot(names(intervalDataWeekday),intervalDataWeekday, type= "l", main = "Weekday",xlab = "Time Interval", ylab = "Steps" )
plot(names(intervalDataWeekend),intervalDataWeekend, type= "l", main = "Weekend", xlab = "Time Interval", ylab = "Steps")

```